```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

## Dictionaries

A dictionary is a way of storing data, associating each item in the dictionary with a name. The common example used to conceptualize this is a phone book - each name is associated with a phone number. 

We can define a dictionary using the following syntax:

```{r, engine='python', eval=FALSE}
dictionaryName = {<key1> : <value1>, 
                  <key2> : <value2>}
```

The `keys` in a dictionary are conventionally a string or an integer. The `values` can be most data types (lists, integers, floats, etc.).

A simple creation of a dictionary: 

```{r, engine='python'}
codons = {'Phe' : ['UUG'], 
          'Leu' : ['CUC']}
          
print(codons)
```

### Looking up values

We can look up the values in a dictionary as such: 

```{r, engine='python'}
codons['Phe']
```

We can also do this by storing the key name as a variable: 

```{r, engine='python'}
myKey = 'Phe'
codons[myKey]
```

### Adding to a dictionary 

We add to a dictionary as such: 

```{r, engine='python', eval=FALSE}
<dictionary name>[<key>] = <value> 
```

If the key already exists, then you will **overwrite** the existing value: 

```{r, engine='python'}
print(codons['Leu'])
codons['Leu'] = ['CUU']
print(codons['Leu'])
```

If the key does not already exist in your dictionary, you will add a new key-value pair: 

```{r, engine='python'}
codons['Ala'] = ['GCC']
print(codons)
```

Likewise, we can use this syntax to modify existing values: 

```{r, engine='python'}
print(codons['Leu'])
codons['Leu'].append('CUG')
print(codons['Leu'])
```

This takes the codon value for `Leu`, which is the list `['CUU']`. Then, we append the value `'CUG'`.

### Looping through dictionaries

We can get the `keys` of a dictionary using the method: `.keys()`:

```{r, engine='python'}
for i in codons.keys():
  print(i)
```

We can do the same for value using `.values()`:

```{r, engine='python'}
for i in codons.values():
  print(i)
```

We can use `.items()` to iterate through the keys and values together:

```{r, engine='python'}
for key, value in codons.items():
  print(key, value)
```

Lastly, if we want the data returned by `.keys()`, `.values()`, and `.items()` stored as list, we can do: 

```{r, engine='python'}
dictionary_keys_list = list(codons.keys())
print(dictionary_keys_list)
```






