```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

## While Loops 

Like a `for` loop, a `while` loop is a way of repeatedly performing an action. However, whereas a `for` loop runs a code block a set number of times, a `while` loop runs the block until a specific condition is met. 

A `while` loop is structured like this:

```
while <condition>:
    {do something}
```

The condition can be any expression that can be evaluated as a Boolean (`True` or `False`) - in particular the comparison operators (`==`, `<`, `>`, etc.) are often used here. For example: 

```{r, engine='python', eval=FALSE}
myString = "A"

while len(myString) <= 10:
  myString = myString + "AAAA"
  print(myString)
```

### Infinite Loops

Because a `while` loop runs an undetermined number times, a common error is an infinite loop, or a `while` loop that never stops running because the condition for termination is never met. Consider the following block of code: 

```{r, engine='python', eval=FALSE}
myNumber = 1 

while myNumber > 0:
  myNumber = myNumber + 1
  print(myNumber)
```

Here, the number starts at 1 (which is greater than 0) and only ever grows larger. The statement `myNumber > 0` will remain true with every iteration of the loop, and the loop will never terminate, preventing your code from fully executing.














